---
title: "01_TempMunge"
author: "Simon Topp"
date: "07/08/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(purrr)
library(furrr)
library(data.table)
library(feather)
library(sf)
library(ggpmisc)
library(ggpubr)
library(Hmisc)
library(Metrics)

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r eval = F, include = F}
## Read in whatever path you saved the raw pull data to
files <- list.files('data/EE_TempPull', full.names = T)
files <- files[file.size(files) > 1] ## 6 WRS tile returned no values

## Munge it a little, basically zero clouds/shadow/ice and
## at least 5 pixels
munger <- function(file){
  df <- read_csv(file) %>%
    filter(!is.na(temp),
           cScore_clouds == 0,
           pCount_water > 5) %>%
    mutate(LandsatID = map_chr(`system:index`, ~str_split(.,'_0000')[[1]][1])) %>%
  select(-`system:index`)
  return(df)
}

munged <- files %>% map_dfr(munger)

munged <- munged %>% mutate(temp = 0.00341802*temp + 149 - 273.15,
                                  temp_qa = temp_qa*0.01,
                                  lake_mix_layer_temperature = lake_mix_layer_temperature - 273.15) %>%
  rename(temp_e5_mix = lake_mix_layer_temperature, temp_e5_full = lake_total_layer_temperature, 
         temp_ls = temp, temp_ls_qa = temp_qa, date = DATE_ACQUIRED, sat = SPACECRAFT_ID, 
         scene_cloud_cover = CLOUD_COVER)

write_feather(munged,'data/out/landsat_temps.feather')
```

```{r}
munged <- read_feather('data/out/landsat_temps.feather')
```

## Between Landsat 7 and 8 we have `r nrow(munged)` temperature observations.

```{r}
obs <- read_csv('data/in/lake_surface_temp_obs.csv') %>%
  rename(date = Date)

same_day <- obs %>% inner_join(munged) %>% mutate(join = 'same_day')

day_plus_one <- obs %>% inner_join(munged %>% mutate(date = date + 1)) %>% mutate(join = 'day_plus_1')

day_minus_one <- obs %>% inner_join(munged %>% mutate(date = date -1)) %>% mutate(join = 'day_minus_1')

matchups <- bind_rows(same_day, day_plus_one, day_minus_one) %>%
  mutate(resid = wtemp_obs - temp_ls,
         abs_error = abs(resid),
         e5_flag = abs(wtemp_obs - temp_e5_mix))

matchups_filt <- matchups %>%
  filter(temp_ls_qa < 3,
         scene_cloud_cover < 20,
         distance > 90,
         temp_ls >=0)

rm(same_day, day_plus_one, day_minus_one)
```

## Of those, we have `r nrow(matchups)` coincident observations of RS and *in situ* temp (matchups).

Let's look at some *inclusive* metrics. General filters here are:

-   +/- 1 day matchups

-   0 clouds/ice/cloud shadow over the 'deepest point'

```{r}
ggplot(matchups, aes(x = wtemp_obs, y = temp_ls)) + geom_hex() +
  scale_fill_viridis_c(trans = 'log10') +
  geom_smooth(method = 'lm') +
  stat_poly_eq(formula = y~x, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE) +
  labs(x = 'Observed Temps', y = 'Landsat Temps', title = 'Landsat and Observed')

matchups %>% summarise(rmse = rmse(wtemp_obs, temp_ls),
                       mae = mae(wtemp_obs,temp_ls),
                       smape = smape(wtemp_obs,temp_ls),
                       bias = bias(wtemp_obs, temp_ls)) %>%
  mutate(across(everything(), ~round(.,4))) %>%
  kableExtra::kable() %>%
  kableExtra::kable_styling()
  
## One quick look at Era5, it's not great
ggplot(matchups, aes(x = wtemp_obs, y = temp_e5_mix)) + geom_hex() +
  scale_fill_viridis_c(trans = 'log10') +
  geom_smooth(method = 'lm', aes(color = 'Best Fit')) +
  geom_abline(aes(slope=1, intercept=0,color = '1:1')) +
  stat_poly_eq(formula = y~x, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE) +
  labs(x = 'Observed Temps', y = 'Era5 Mixed Layer Temps', 'Era5 and Observed', titel='blue=Best Fit', 'Red=1:1')


matchups %>% filter(!is.na(temp_e5_mix)) %>%
  summarise(rmse = rmse(wtemp_obs, temp_e5_mix),
                       mae = mae(wtemp_obs,temp_e5_mix),
                       smape = smape(wtemp_obs,temp_e5_mix),
                       bias = bias(wtemp_obs, temp_e5_mix)) %>%
  mutate(across(everything(), ~round(.,4))) %>%
  kableExtra::kable() %>%
  kableExtra::kable_styling()
```

## Now the same but with *strict* filters. After qc filters we have `r nrow(matchups_filt)` matchups.

General filters here are:

-   +/- 1 day matchups

-   0 clouds/ice/cloud shadow over the 'deepest point'

-   Landsat uncertainty \< 3 degrees

-   Scene cloud cover \< 20% - Distance to shore \> 90m

```{r}
matchups_filt <- matchups %>%
  filter(temp_ls_qa < 3,
         scene_cloud_cover < 10,
         distance > 200,
         temp_ls >=0,
         wtemp_obs >=0,
         e5_flag < 5,
         join != 'day_minus_1'
         )

plotter <- function(var){
  matchups %>%
    mutate(bins= cut_number(get(var), 5)) %>%
    ggplot(aes(x = bins, y = abs_error)) +
    geom_violin(draw_quantiles = c(.25,.5,.75)) +
    geom_hline(aes(yintercept = 0), color = 'red')+
    labs(x = var)
}

plotter('areakm')
plotter('scene_cloud_cover')
plotter('e5_flag')
plotter('distance')
plotter('temp_ls_qa')

matchups_filt <- matchups %>%
  filter(areakm > .2,
         temp_ls_qa < 3,
         scene_cloud_cover < 20,
         distance > 250,
         temp_ls >= 0,
         wtemp_obs >=0,
         e5_flag < 5,
         join != 'day_minus_1'
         )

ggplot(matchups_filt, aes(x = wtemp_obs, y = temp_ls)) + geom_hex() +
  scale_fill_viridis_c(trans = 'log10') +
  geom_smooth(method = 'lm') +
  stat_poly_eq(formula = y~x, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE) +
  labs(x = 'Observed Temps', y = 'Landsat Temps', title = 'Landsat and Observed')

matchups_filt %>% summarise(rmse = rmse(wtemp_obs, temp_ls),
                       mae = mae(wtemp_obs,temp_ls),
                       smape = smape(wtemp_obs,temp_ls),
                       bias = bias(wtemp_obs, temp_ls)) %>%
  mutate(across(everything(), ~round(.,4))) %>%
  kableExtra::kable() 

check <- matchups_filt %>% mutate(se = (wtemp_obs - temp_ls)^2)
```

## Finally, just big lakes where we would expect to be most confident (distance to shore \> 1000m)

```{r}
ggplot(matchups_filt %>% filter(distance > 1000), aes(x = wtemp_obs, y = temp_e5_mix)) + geom_hex() +
  scale_fill_viridis_c(trans = 'log10') +
  geom_smooth(method = 'lm') +
  stat_poly_eq(formula = y~x, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE) +
  labs(x = 'Observed Temps', y = 'Landsat Temps', title = 'Landsat and Observed')

matchups_filt %>% filter(distance > 1000) %>%
  summarise(rmse = rmse(wtemp_obs, temp_ls),
                   mae = mae(wtemp_obs,temp_ls),
                   smape = smape(wtemp_obs,temp_ls),
                   bias = bias(wtemp_obs, temp_ls)) %>%
  mutate(across(everything(), ~round(.,4))) %>%
  kableExtra::kable() %>%
  kableExtra::kable_styling()
```

## Sliced and diced a couple ways

```{r}
lake_data <- read_csv('data/in/lake_metadata.csv')

matchups <- matchups %>% left_join(lake_data) %>% 
  mutate(residual = wtemp_obs - temp_ls,
         abs_error = abs(residual))

ggplot(matchups, aes(x = residual, fill = sat)) +
  geom_density(alpha = .4) +
  geom_vline(aes(xintercept = 0), color = 'red') +
  labs(title = 'Residuals across sensors')

ggplot(matchups, aes(x = residual, fill = join)) +
  geom_density(alpha = .4) +
  geom_vline(aes(xintercept = 0), color = 'red') +
  labs(title = 'Residuals across sensors')

matchups %>%
  mutate(dist_group = cut_number(distance, 6)) %>%
  ggplot(aes(x = dist_group, y = residual)) +
  geom_violin(draw_quantiles = .5) +
  labs(title = 'Residuals binned by distance to shore')


## Spatially
matchups.sf <- matchups %>% st_as_sf(coords = c('lake_lon_deg','lake_lat_deg'), crs = 4326) %>% st_transform(5070)

usa <- maps::map('usa', plot = F) %>% st_as_sf() %>% st_transform(5070) 

grid <- st_make_grid(usa, cellsize = c(200000,200000), square = F) %>% st_as_sf() %>% mutate(ID = row_number())

rmse.sf <- matchups.sf %>% st_join(grid) %>% st_set_geometry(NULL) %>%
  group_by(ID) %>%
  summarise(rmse = rmse(wtemp_obs, temp_ls),
            matchup.count = n()) %>%
  inner_join(grid) %>%
  st_as_sf() 

ggplot(rmse.sf) + geom_sf(aes(fill = rmse)) +
  scale_fill_viridis_c() +
  geom_sf(data = usa, fill = 'transparent', color = 'black') +
  theme_minimal() +
  labs(title = 'Spatial Error Distribution')

ggplot(rmse.sf) + geom_sf(aes(fill = matchup.count)) +
  scale_fill_viridis_c(trans='log10') +
  geom_sf(data = usa, fill = 'transparent', color = 'black') +
  theme_minimal() +
  labs(title = 'Matchup Count')

```

## Take a look at within scene variance.
```{r}
summary_in <- matchups_filt

scene_sd <- summary_in %>% group_by(LandsatID) %>%
  summarise(count = n(),
            resid_sd = sd(resid)) %>%
  filter(count >5)

overall_sd <- summary_in %>% filter(LandsatID %in% scene_sd$LandsatID) %>%
  summarise(overall = sd(resid)) %>% .[[1]]

ggplot(scene_sd, aes(x = resid_sd)) + geom_histogram() + geom_vline(aes(xintercept = overall_sd, color = 'Overall residual SD'))

scene_counts <- dplyr::count(summary_in, LandsatID) %>% filter(n > 5)

scene_bias <- summary_in %>% 
  filter(LandsatID %in% scene_counts$LandsatID) %>%
  group_by(LandsatID) %>%
  summarise(bias = bias(wtemp_obs, temp_ls))

scene_stats <- summary_in %>% 
  filter(LandsatID %in% scene_counts$LandsatID) %>%
  left_join(scene_bias) %>%
  mutate(temp_no_bias = temp_ls+bias) %>%
  group_by(LandsatID) %>%
  summarise(rmse_bias_removed = rmse(wtemp_obs, temp_no_bias),
            scene_bias = mean(bias))

overall_bias = summary_in %>% filter(LandsatID %in% scene_counts$LandsatID) %>%
  summarise(bias = bias(wtemp_obs, temp_ls)) %>%
  .[[1]]

overall_rmse = summary_in %>% filter(LandsatID %in% scene_counts$LandsatID) %>%
  mutate(temp_ls = temp_ls + overall_bias) %>%
  summarise(rmse = rmse(wtemp_obs,temp_ls))
rmses <- tibble(overall_rmse_bias_removed = overall_rmse[[1]], median_scene_rmse_bias_removed = median(scene_stats$rmse_bias_removed)) %>%
  pivot_longer(everything(), names_to = "RMSE_Summaries")

ggplot(scene_stats, aes(x = rmse_bias_removed)) + geom_histogram() + 
  geom_vline(data = rmses, aes(xintercept = value, color = RMSE_Summaries)) +
  labs(x = "Distribution of image specific rmse after removing image bias")
```


```{r}
mendota <- read_csv('data/in/Mendota_Era5Land_1984_2021.csv') 
mendotaObs <- obs %>% filter(site_id == "nhdhr_143249470")

mendota <- mendota %>% mutate(lake_mix_layer_temperature = lake_mix_layer_temperature - 273.15,
                              date = as.Date(date)) %>%
  inner_join(mendotaObs)

library(plotly)

ggplotly(
  ggplot(mendota, aes(x=date)) +
    geom_point(aes(y = lake_mix_layer_temperature, color='Era5')) +
    geom_point(aes(y=wtemp_obs, color = 'Obs')) +
    geom_line(aes(y = lake_mix_layer_temperature, color='Era5')) +
    geom_line(aes(y=wtemp_obs, color = 'Obs')))

ggplot(mendota, aes(x=lake_mix_layer_temperature)) + geom_histogram()

ggplot


mendota %>% rename(temp_e5_mix = lake_mix_layer_temperature) %>%
  summarise(rmse = rmse(wtemp_obs, temp_e5_mix),
                       mae = mae(wtemp_obs,temp_e5_mix),
                       smape = smape(wtemp_obs,temp_e5_mix),
                       bias = bias(wtemp_obs, temp_e5_mix)) %>%
  mutate(across(everything(), ~round(.,4))) %>%
  kableExtra::kable() %>%
  kableExtra::kable_styling()

mendota %>% rename(temp_e5_mix = lake_mix_layer_temperature) %>%
ggplot(.,aes(x = wtemp_obs, y = temp_e5_mix)) + geom_hex() +
  scale_fill_viridis_c(trans = 'log10') +
  geom_smooth(method = 'lm', aes(color = 'Best Fit')) +
  geom_abline(aes(slope=1, intercept=0,color = '1:1')) +
  stat_poly_eq(formula = y~x, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE) +
  labs(x = 'Observed Temps', y = 'Era5 Mixed Layer Temps', 'Era5 and Observed', titel='blue=Best Fit', 'Red=1:1')
```

